# Linear regression

## Background

- Simplest regression model: $y=a+bx+error$
- Additional predictors: : $y=\beta_0+\beta_1x_1+...+\beta_kx_k+error$
- Nonlinear models: $\log y=a+b \log x+error$
- Nonadditive models: $y=\beta_0+\beta_1x_1+\beta_2x_2+\beta_3x_1x_2+error$
- Generalized linear models: Non-normal additive errors with a function that "links" the linear predictor with the outcome.

## Simple regression with fake data

Same as Chapter \@ref(intro)

```{r warning=FALSE,message=FALSE}
x=1:20
n=length(x)
a=0.2
b=0.3
sigma=0.5
y=a + b*x + sigma*rnorm(n)
```

```{r warning=FALSE,message=FALSE}
fit=lm(y~x)
summary(fit)
```

```{r echo=FALSE}
kable(confint(fit), caption = '95% CI for intercept and slope',
  booktabs = TRUE)
```

*Questions*

1. Are estimates consistent with true values?
2. What can you tell about the uncertainty surrounding the coefficients?

We can also plot the data and fitter regression line (Figure \@ref(fig:fig1)).

```{r fig1, warning=FALSE,message=FALSE, fig.cap="Observed and fitted"}
plot(x, y, main="Data and fitted regression line")
a_hat=coef(fit)[1]
b_hat=coef(fit)[2]
abline(a_hat, b_hat)
```

## Interpretation of coefficients and model diagnostics

*Linear regression assumptions*

1. Normality of residuals.
2. Linear relationship.
3. Constant variance.
4. $y_i$ are independent.


Let's start using real data. Below we can see a brief description about the dataset used in this example.

```{r warning=FALSE,message=FALSE,eval=FALSE}
install.packages("datarium") # Install to access the dataset
?marketing #description of the dataset
library(datarium)
head(marketing)
```

```{r warning=FALSE,message=FALSE,echo=FALSE}
library(datarium)
kable(head(marketing), booktabs = TRUE)
```

We build a model to predict sales on the basis of advertising budget spent in youtube **and** facebook. It is important to notice the nonadditive model used here and the implications of interaction terms.

```{r warning=FALSE,message=FALSE,eval}
fit2=lm(sales ~ youtube*facebook, data = marketing)
summary(fit2)
```

Also we plot some regression diagnostics.

```{r fig2, warning=FALSE,message=FALSE}
plot(fit2)
```