--- 
title: "Tutorial notes for Biostatistics II"
author: "Juan Pablo Diaz Martinez"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: book.bib
biblio-style: apalike
link-citations: yes
description: This is a minimal example of using the bookdown package to write a book.
  The output format for this example is bookdown::gitbook.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```


# Common pitfalls in statistical analysis

## p-values

+ The p-value is the probability of obtaining test results at least as extreme as the result actually observed, under the assumption that the null hypothesis is correct.

**Example**

Suppose we want to compare fasting serum-cholesterol levels among recent Asian immigrants to the United States with typical levels found in the general U.S. population. Suppose we assume cholesterol levels in women ages 21−40 in the United States are approximately normally distributed with mean 190 mg/dL. It is unknown whether cholesterol levels among recent Asian immigrants are higher or lower than those in the general U.S. population. Let’s assume that levels among recent female Asian immigrants are normally distributed with unknown mean $\mu$. Hence we wish to test

$$
\begin{aligned}
 H_0 &: \mu=\mu_0 \\
 H_1 &: \mu \neq \mu_0 
\end{aligned}
$$

Blod tests are performed on 100 female Asian immigrants ages 21−40, and the mean level (x) is 181.52 mg/dL with standard error = 40 mg/dL. What can we conclude on the basis of this evidence?

First, the test statistic would be:

$$
\begin{aligned}
 t &= (\bar{x}-\mu_0)/(s/\sqrt{n}) \\
 t &= -8.48/4 = -2.12
\end{aligned}
$$

Under the null, this test statistic follows a t-distribution with $n-1$ degrees of freedom, i.e,

$$
t \sim t(n-1)
$$

Mathematically, for symmetric distributions (such as the t-distribution), the p-value for two-sided tests is define as:

$$
\begin{aligned}
p &=2*\Pr[|T| \geq |t| \space | \space H_0] \\
p &=2*\Pr[|T| \geq 2.12 \space | \space H_0]=0.37
\end{aligned}
$$

where $T$ is the distribution under the null and $t$

```{r}
2*pt(2.12,df=99,lower.tail = F)
```

**Missconceptions behind p-values**

1. The p-value is not the probability that the null hypothesis is true, or the probability that the alternative hypothesis is false
2. The p-value is not the probability that the observed effects were produced by random chance alone
3. The 0.05 significance level is merely a convention
4. The p-value does not indicate the size or importance of the observed effect

## Table 1 

+ It is recommended to not report p-values when comparing summary statistics between two or more groups because:
  1. Table 1 usually describes the study sample
  2. Valid p-values cannot be drawn without knowing, not just what was done with the existing data, but what the choices in data coding, exclusion, and analysis would have been, had the data been different
  3. Being significant does not imply anything in a regression context
+ One recommendation is to report the standardized mean difference (SMD). The p-values are a function of a sample size, meaning that with a large sample most likelu we will find a significal difference.
  
## Variables

+ Do not test for normality. Majority of parametric tests assumptions are based on the  *mean* or the residuals (linear regression model)
+ Report missing values


